---
author: "Harly Durbin"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
    code_folding: hide
params:
  rundate: "190812"
  r1: "3"
  r2: "2"
title: "`r glue::glue('Correlation between breeding values (region {params$r1} vs. region {params$r2})')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(purrr)
library(ggplot2)
library(tidylog)
library(glue)
library(rlang)

source(here::here("source_functions/gather_solutions.R"))

```

```{r}
round_any <- function(x, accuracy, f = round){f(x/accuracy)*accuracy}

```


```{r}
region_key <-
  tribble(~num, ~abbrv, ~desc,
          2, "SE", "Southeast",
          8, "FB", "Fescue Belt",
          3, "HP", "High Plains", 
          5, "AP", "Arid Prairie",
          7, "FM", "Forested Mountains", 
          1, "D", "Desert",
          9, "UMWNE", "Upper Midwest & Northeast")
```


```{r}

r1_abbrv <- 
  region_key %>% 
  filter(num == params$r1) %>% 
  pull(abbrv)
  
r1_desc <-  
  region_key %>% 
  filter(num == params$r1) %>% 
  pull(desc)
 

r2_abbrv <- 
  region_key %>% 
  filter(num == params$r2) %>% 
  pull(abbrv)


r2_desc <-  
  region_key %>% 
  filter(num == params$r2) %>% 
  pull(desc)

direc <- glue("{params$rundate}_{params$r1}v{params$r2}")

```


# Import/setup

```{r}

sol_univ <-
  gather_univ(
    rundate = params$rundate,
    r1 = params$r1,
    r2 = params$r2, 
    growth_trait = "ww"
  )

```


```{r}

sol_biv <-
  gather_biv(
    rundate = params$rundate,
    r1 = params$r1,
    r2 = params$r2, 
    growth_trait = "ww"
  )

```

```{r full bvs}


sol_full <-
read_rds(here::here("data/derived_data/animal_regions.rds")) %>%
  filter(id_new %in% sol_biv$id_new) %>% 
  filter(trait == "ww") %>% 
  filter(var %in% c("bv_sol", "mat_sol")) %>% 
  mutate(model = "full") %>% 
  select(
    id_new,
    analysis_region = region, 
    effect = var,
    solution = value,
    model
         ) 


```

```{r}

dat_ped <-
  list(glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/{r1_abbrv}/data.{r1_abbrv}.txt"), glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/{r2_abbrv}/data.{r2_abbrv}.txt")) %>% 
  set_names(params$r1, params$r2) %>% 
  map_dfr(~ read_table2(
    here::here(.x), 
    col_names = c("id_new", "cg_new", "weight")
  ),
  .id = "region") %>% 
  left_join(
      list(glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/{r1_abbrv}/ped.{r1_abbrv}.txt"), glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/{r2_abbrv}/ped.{r2_abbrv}.txt")) %>% 
        set_names(params$r1, params$r2) %>% 
        map_dfr(~ read_table2(
          here::here(.x), 
          col_names = c("id_new", "sire_id", "dam_id")
          ),
          .id = "region")
      )

```

# `r glue("Plot bivariate {r1_desc} breeding values against {r2_desc} breeding values")`

## Weaning weight


```{r plotbivww, fig.width=10, fig.height=8}

lowlim <- round_any(min(sol_biv[sol_biv$effect == "bv_sol", "solution"]), 5)
hilim <- round_any(max(sol_biv[sol_biv$effect == "bv_sol", "solution"]), 5)



sol_biv %>%
  filter(effect == "bv_sol") %>% 
  tidyr::pivot_wider(
    names_from = analysis_region,
    names_prefix = "region",
    values_from = "solution"
    ) %>%
  ggplot(aes(
    x = !!sym(glue("region{params$r1}")),
    y = !!sym(glue("region{params$r2}"))
    )) +
  geom_hex(aes(fill = stat(count)),
           bins = 65) +
  viridis::scale_fill_viridis(option = "inferno",
                              direction = 1,
                              begin = 0.2,
                              labels = scales::comma
                              ) +
  geom_abline(slope = 1,
              size = 1,
              linetype = "twodash") +
  lims(
    x = c(lowlim, hilim),
    y = c(lowlim, hilim)
       ) +
  theme_classic() +
  theme(
    plot.title = element_text(size = 28,
                              face = "italic"),
    plot.subtitle = element_text(size = 20,
                                 margin = margin(
                                   t = 0,
                                   r = 0,
                                   b = 13,
                                   l = 0
                                 )),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 14),
    axis.title = element_text(size = 22),
    axis.title.y = element_text(margin = margin(
      t = 0,
      r = 13,
      b = 0,
      l = 0
    )),
    axis.title.x = element_text(margin = margin(
      t = 13,
      r = 0,
      b = 0,
      l = 0
    )),
    axis.text = element_text(size = 16)
  ) +
  labs(x = glue("{r1_desc} BV"),
       y = glue("{r2_desc} BV"),
       fill = "Count",
       title = glue("Estimated breeding values: {r1_desc} \nvs. {r2_desc}")
       )

```

```{r plotbivww2, fig.width=10, fig.height=8}

lowlim <- round_any(min(sol_biv[sol_biv$effect == "bv_sol", "solution"]), 5)
hilim <- round_any(max(sol_biv[sol_biv$effect == "bv_sol", "solution"]), 5)

dat_ped %>%
  group_by(sire_id, region) %>%
  tally() %>%
  ungroup() %>%
  tidyr::pivot_wider(
    id_cols = "sire_id",
    names_prefix = "n_",
    names_from = region,
    values_from = n
  ) %>%
  filter(!!sym(glue("n_{params$r1}")) >= 10) %>%
  filter(!!sym(glue("n_{params$r2}")) >= 10) %>%
  left_join(sol_biv %>%
              filter(effect == "bv_sol"),
            by = c("sire_id" = "id_new")) %>%
  tidyr::pivot_wider(
    id_cols = "sire_id",
    names_prefix = "bivariate",
    names_from = analysis_region,
    values_from = solution
  ) %>%
  mutate(change = !!sym(glue("bivariate{params$r1}")) -!!sym(glue("bivariate{params$r2}"))) %>%
  ggplot(aes(x = !!sym(glue(
    "bivariate{params$r1}"
  )),
  y = !!sym(glue(
    "bivariate{params$r2}"
  )))) +
  geom_point(aes(color = change), size = 3) +
  viridis::scale_color_viridis(
    option = "inferno",
    direction = 1,
    begin = 0.2,
    labels = scales::comma
  ) +
  geom_abline(slope = 1,
              size = 1,
              linetype = "twodash") +
  lims(x = c(lowlim, hilim),
       y = c(lowlim, hilim)) +
  theme_classic() +
  theme(
    plot.title = element_text(size = 28,
                              face = "italic"),
    plot.subtitle = element_text(size = 20,
                                 margin = margin(
                                   t = 0,
                                   r = 0,
                                   b = 13,
                                   l = 0
                                 )),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 14),
    axis.title = element_text(size = 22),
    axis.title.y = element_text(margin = margin(
      t = 0,
      r = 13,
      b = 0,
      l = 0
    )),
    axis.title.x = element_text(margin = margin(
      t = 13,
      r = 0,
      b = 0,
      l = 0
    )),
    axis.text = element_text(size = 16)
  ) +
  labs(
    x = glue("{r1_desc} weaning weight BV"),
    y = glue("{r2_desc} weaning weight BV"),
    title = glue("Estimated weaning weight breeding\nvalues: {r1_desc} vs. {r2_desc}"),
    subtitle = "(Sires with at least 10 analyzed calves in both regions)",
    color = "Change"
  )


```

## Milk 

```{r plotbivmilk, fig.width=10, fig.height=8}

lowlim <- round_any(min(sol_biv[sol_biv$effect == "mat_sol", "solution"]), 5)
hilim <- round_any(max(sol_biv[sol_biv$effect == "mat_sol", "solution"]), 5)

sol_biv %>%
  filter(effect == "mat_sol") %>% 
  tidyr::pivot_wider(
    names_from = analysis_region,
    names_prefix = "region",
    values_from = "solution"
    ) %>%
  ggplot(aes(
    x = !!sym(glue("region{params$r1}")),
    y = !!sym(glue("region{params$r2}"))
    )) +
  geom_hex(aes(fill = stat(count)),
           bins = 65) +
  viridis::scale_fill_viridis(option = "inferno",
                              direction = 1,
                              begin = 0.2,
                              labels = scales::comma
                              ) +
  geom_abline(slope = 1,
              size = 1,
              linetype = "twodash") +
  lims(
    x = c(lowlim, hilim),
    y = c(lowlim, hilim)
       ) +
  theme_classic() +
  theme(
    plot.title = element_text(size = 28,
                              face = "italic"),
    plot.subtitle = element_text(size = 20,
                                 margin = margin(
                                   t = 0,
                                   r = 0,
                                   b = 13,
                                   l = 0
                                 )),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 14),
    axis.title = element_text(size = 22),
    axis.title.y = element_text(margin = margin(
      t = 0,
      r = 13,
      b = 0,
      l = 0
    )),
    axis.title.x = element_text(margin = margin(
      t = 13,
      r = 0,
      b = 0,
      l = 0
    )),
    axis.text = element_text(size = 16)
  ) +
  labs(x = glue("{r1_desc} milk BV"),
       y = glue("{r2_desc} milk BV"),
       fill = "Count",
       title = glue("Estimated milk breeding values:\n{r1_desc} vs. {r2_desc}")
       )

```

```{r, fig.width=10, fig.height=8}

lowlim <- round_any(min(sol_biv[sol_biv$effect == "mat_sol", "solution"]), 5)
hilim <- round_any(max(sol_biv[sol_biv$effect == "mat_sol", "solution"]), 5)

dat_ped %>%
  group_by(sire_id, region) %>%
  tally() %>%
  ungroup() %>%
  tidyr::pivot_wider(
    id_cols = "sire_id",
    names_prefix = "n_",
    names_from = region,
    values_from = n
  ) %>%
  filter(!!sym(glue("n_{params$r1}")) >= 10) %>%
  filter(!!sym(glue("n_{params$r2}")) >= 10) %>%
  left_join(sol_biv %>%
              filter(effect == "mat_sol"),
            by = c("sire_id" = "id_new")) %>%
  tidyr::pivot_wider(
    id_cols = "sire_id",
    names_prefix = "bivariate",
    names_from = analysis_region,
    values_from = solution
  ) %>%
  mutate(change = !!sym(glue("bivariate{params$r1}")) -!!sym(glue("bivariate{params$r2}"))) %>%
  ggplot(aes(x = !!sym(glue(
    "bivariate{params$r1}"
  )),
  y = !!sym(glue(
    "bivariate{params$r2}"
  )))) +
  geom_point(aes(color = change), size = 3) +
  viridis::scale_color_viridis(
    option = "inferno",
    direction = 1,
    begin = 0.2,
    labels = scales::comma
  ) +
  geom_abline(slope = 1,
              size = 1,
              linetype = "twodash") +
  lims(x = c(lowlim, hilim),
       y = c(lowlim, hilim)) +
  theme_classic() +
  theme(
    plot.title = element_text(size = 28,
                              face = "italic"),
    plot.subtitle = element_text(size = 20,
                                 margin = margin(
                                   t = 0,
                                   r = 0,
                                   b = 13,
                                   l = 0
                                 )),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 14),
    axis.title = element_text(size = 22),
    axis.title.y = element_text(margin = margin(
      t = 0,
      r = 13,
      b = 0,
      l = 0
    )),
    axis.title.x = element_text(margin = margin(
      t = 13,
      r = 0,
      b = 0,
      l = 0
    )),
    axis.text = element_text(size = 16)
  ) +
  labs(
    x = glue("{r1_desc} milk BV"),
    y = glue("{r2_desc} milk BV"),
    title = glue("Estimated milk breeding values:\n{r1_desc} vs. {r2_desc}"),
    subtitle = "(Sires with at least 10 calves in both regions)",
    color = "Change"
  )


```


# Correlation between univariate and bivariate BVs

```{r}
sol_full
```

```{r}
sol_full %>% 
  distinct(effect)
```


```{r}

corrs <-
  bind_rows(sol_biv, sol_univ, sol_full %>% 
              select(-analysis_region)) %>% 
  filter(effect %in% c("bv_sol", "mat_sol")) %>% 
  mutate(model = as.character(glue("{model}{analysis_region}", .na = ""))) %>% 
  arrange(id_new) %>% 
  tidyr::pivot_wider(id_cols = c("id_new", "effect"),
                     names_from = model,
                     values_from = solution) %>% 
  select(id_new, effect, starts_with("uni"), starts_with("biv"), full)

```

## All animals

### Weaning weight direct

```{r}
corrs %>% 
  filter(effect == "bv_sol") %>% 
  select(-id_new, -effect) %>% 
  set_names(c(glue("{r1_desc} single-trait"), glue("{r2_desc} single-trait"), glue("{r1_desc} multi-trait"), glue("{r2_desc} multi-trait"), "Full run")) %>% 
  corrr::correlate(diagonal = 1) 
  

```

### Maternal milk 

```{r}
corrs %>% 
  filter(effect == "mat_sol") %>% 
  select(-id_new, -effect) %>% 
  set_names(c(glue("{r1_desc} single-trait"), glue("{r2_desc} single-trait"), glue("{r1_desc} multi-trait"), glue("{r2_desc} multi-trait"), "Full run")) %>% 
  corrr::correlate(diagonal = 1) 
  

```


## Does correlation between breeding values change when considering only sires with calves in both regions? (n = `r dat_ped %>% group_by(sire_id, region) %>% tally() %>% filter(n_distinct(region) > 1) %>% ungroup() %>% n_distinct(.$sire_id) %>% scales::comma(.)`)

### Weaning weight direct

```{r}
dat_ped %>%
  group_by(sire_id, region) %>%
  tally() %>% 
  filter(n_distinct(region) > 1) %>% 
  ungroup() %>% 
  select(id_new = sire_id) %>% 
  left_join(corrs) %>%
  filter(effect == "bv_sol") %>% 
  select(-id_new, -effect) %>% 
  set_names(c(glue("{r1_desc} single-trait"), glue("{r2_desc} single-trait"), glue("{r1_desc} multi-trait"), glue("{r2_desc} multi-trait"), "Full run")) %>% 
  corrr::correlate(diagonal = 1) 
```

### Maternal milk

```{r}
dat_ped %>%
  group_by(sire_id, region) %>%
  tally() %>% 
  filter(n_distinct(region) > 1) %>% 
  ungroup() %>% 
  select(id_new = sire_id) %>% 
  left_join(corrs) %>%
  filter(effect == "mat_sol") %>% 
  select(-id_new, -effect) %>% 
  set_names(c(glue("{r1_desc} single-trait"), glue("{r2_desc} single-trait"), glue("{r1_desc} multi-trait"), glue("{r2_desc} multi-trait"), "Full run")) %>% 
  corrr::correlate(diagonal = 1) 
```


---

```{r, eval = FALSE}
## ...and their own record (n = `r corrs %>% filter_all(all_vars(!is.na(.))) %>% filter(id_new %in% dat$id_new) %>% n_distinct(.$id_new) %>% scales::comma(.)`)
corrs %>% 
  filter_all(all_vars(!is.na(.))) %>% 
  filter(id_new %in% dat$id_new) %>% 
  select(-id_new) %>% 
  set_names(c(glue("{r1_desc} single-trait"), glue("{r2_desc} single-trait"), glue("{r1_desc} multi-trait"), glue("{r2_desc} multi-trait"), "Full run")) %>% 
  corrr::correlate(diagonal = 1) 
```



```{r, eval = FALSE}
# Re-ranking
bulls <-
  bind_rows(sol_biv, sol_univ, sol_full %>%
            select(-analysis_region)) %>%
  filter(effect == "bv_sol") %>%
  mutate(analysis = as.character(glue("{analysis}{analysis_region}", .na = ""))) %>%
  arrange(id_new) %>%
  tidyr::pivot_wider(id_cols = "id_new",
                     names_from = analysis,
                     values_from = solution) %>%
  filter_all(all_vars(!is.na(.))) 
```

```{r,eval=FALSE}
bulldiffs <-
  dat %>% 
  mutate(
    region = 
      case_when(
        weight_2 != 0 ~ 2,
        weight_3 != 0 ~ 3
      )
  ) %>% 
  select(id_new, region) %>% 
  left_join(ped %>% 
              select(id_new, sire_id)) %>% 
  group_by(sire_id, region) %>%
  tally() %>%
  ungroup() %>%
  group_by(sire_id) %>%
  filter(n_distinct(region) == 2) %>% 
  ungroup() %>% 
  tidyr::pivot_wider(
    id_cols = c("sire_id"),
    names_from = region,
    values_from = n,
    names_prefix = "n_"
  ) %>% 
  mutate(n_total = n_3 + n_2) %>% 
  rename(id_new = sire_id) %>% 
  left_join(bulls) %>% 
  select(-starts_with("uni")) %>% 
  mutate(SEvHP = bivariate2 - bivariate3,
         HPvSE = bivariate3 - bivariate2) %>% 
  left_join(ped %>% 
              select(id_new, reg))
```


```{r,eval=FALSE}
bulldiffs %>% 
  filter(n_2 >= 50) %>% 
  filter(n_3 >= 50) %>% 
  select(reg, id_new, everything()) %>% 
  arrange(desc(SEvHP))

  
```
