---
title: "Sire summary"
author: "Harly Durbin"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
    code_folding: hide
params:
  rundate: "190812"
  r1: "3"
  r2: "1"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(purrr)
library(ggplot2)
library(tidylog)

region_key <-
  tribble(~num, ~abbrv, ~desc,
          2, "SE", "Southeast",
          8, "FB", "Fescue Belt",
          3, "HP", "High Plains", 
          5, "AP", "Arid Prairie",
          7, "FM", "Forested Mountains", 
          1, "D", "Desert",
          9, "UMWNE", "Upper Midwest & Northeast")

r1_abbrv <- 
  region_key %>% 
  filter(num == params$r1) %>% 
  pull(abbrv)
  
r1_desc <-  
  region_key %>% 
  filter(num == params$r1) %>% 
  pull(desc)
 

r2_abbrv <- 
  region_key %>% 
  filter(num == params$r2) %>% 
  pull(abbrv)


r2_desc <-  
  region_key %>% 
  filter(num == params$r2) %>% 
  pull(desc)

direc <- glue("{params$rundate}_{params$r1}v{params$r2}")

```

```{r, eval=TRUE}
source(here::here("source_functions/ped.R"))

```


# Import/Setup 

```{r}

dat <-
  read_table2(here::here(
    glue("data/f90/{direc}/data.{r1_abbrv}_{r2_abbrv}.txt")),  
              col_names = c("id_new", "cg_new", glue("weight_{params$r1}"), glue("weight_{params$r2}")))


```

```{r, eval=TRUE}

ped_biv <-
  list( glue("data/f90/{direc}/{r1_abbrv}/ped.{r1_abbrv}.txt"),  glue("data/f90/{direc}/{r2_abbrv}/ped.{r2_abbrv}.txt")) %>% 
  rlang::set_names(params$r1, params$r2) %>% 
  map_dfr(~ read_table2(here::here(.x), 
              col_names = c("id_new", "sire_id", "dam_id")),
      .id = "region") %>% 
filter(id_new %in% dat$id_new)


```


# Total sires

```{r}
ped_biv %>%
  group_by(region) %>%
  summarise(`n sires` = n_distinct(sire_id)) %>% 
  rename(Region = region) %>% 
  mutate(`n sires` = scales::comma(`n sires`))
```

## How many sires with progeny in both regions? 

* `r ped_biv %>% group_by(sire_id, region) %>% tally() %>% ungroup() %>% group_by(sire_id) %>% filter(n_distinct(region) == 2) %>% ungroup() %>% distinct(sire_id) %>% nrow() %>% scales::comma()` bulls with progeny in both regions

```{r}

ped_biv %>%
  group_by(sire_id, region) %>%
  tally() %>%
  ungroup() %>% 
  pivot_wider(id_cols = sire_id,
              names_from = region,
              values_from = n, 
              names_prefix = "region_") %>% 
  filter(!is.na(!!sym(glue("region_{params$r1}")))) %>% 
  filter(!is.na(!!sym(glue("region_{params$r2}")))) %>% 
  mutate(total = !!sym(glue("region_{params$r1}")) + !!sym(glue("region_{params$r2}"))) %>% 
  arrange(desc(total))# %>% 
  # DT::datatable(options = list(pageLength = 10), rownames = FALSE, caption = "Sires with calves in both regions")


```


```{r, fig.width=7, fig.height=8}

bullsboth <- ped_biv %>% group_by(sire_id, region) %>% tally() %>% ungroup() %>% group_by(sire_id) %>% filter(n_distinct(region) == 2) %>% ungroup() %>% distinct(sire_id) %>% nrow()

ped_biv %>%
  group_by(sire_id, region) %>%
  tally() %>%
  ungroup() %>%
  group_by(sire_id) %>%
  filter(n_distinct(region) == 2) %>%
  ungroup() %>% 
  ggplot(aes(x = n,
             fill = region)) +
  geom_histogram(bins = 20, 
                 show.legend = FALSE) +
  facet_wrap( ~ region, 
              nrow = 2, 
             labeller = labeller(
               region = c(
                 "1" = "1: Desert",
                 "2" = "2: Southeast",
                 "3" = "3: High Plains",
                 "4" = "4: Rainforest",
                 "5" = "5: Arid Prairie",
                 "6" = "6: Cold Desert",
                 "7" = "7: Forested Mountains",
                 "8" = "8: Fescue Belt",
                 "9" = "9: Upper Midwest & Northeast"
                 )
               )
              ) +
  scale_fill_manual(
    values = c(
      "1" = "tomato2",
      "2" = "darkslategray4",
      "3" = "springgreen3",
      "4" = "brown",
      "5" = "goldenrod1",
      "6" = "gray50",
      "7" = "deeppink3",
      "8" = "gray17",
      "9" = "slateblue2"
    )
  ) +
  labs(title = glue("Distribution of calves, bulls with progeny in\n both regions (n = {bullsboth})"),
       x = "Number of calves",
       y = "Count of bulls") +
  theme_classic() +
    theme(
    plot.title = element_text(
      size = 20,
      face = "italic",
      margin = margin(t = 0, r = 0, b = 13, l = 0)
    ),
    axis.title = element_text(
      size = 16 
    ),
    axis.title.y = element_text(
      margin = margin(t = 0, r = 13, b = 0, l = 0)
    ),
    axis.title.x = element_text(
      margin = margin(t = 13, r = 0, b = 0, l = 0)
    ),
    axis.text = element_text(
      size = 14
    ),
    legend.text = element_text(
      size = 14
    ),
    strip.text.x = element_text(
      size = 14
    )
  )

```


## How many paternal sires with grand-progeny in both regions? 

```{r}

ped3 <-
  ped_biv %>%
  left_join(ped_biv %>%
              distinct(sire_id) %>%
              left_join(
                ped %>%
                  select(id_new,
                         pgs_id = sire_id,
                         pgd_id = dam_id),
                by = c("sire_id" = "id_new")
              ))
```

* `r ped3 %>% group_by(pgs_id, region) %>% tally() %>% ungroup() %>% group_by(pgs_id) %>% filter(n_distinct(region) == 2) %>% ungroup() %>% distinct(pgs_id) %>% nrow() %>% scales::comma()` bulls with grand-progeny in both regions

```{r}
ped3 %>% 
  group_by(pgs_id, region) %>% 
  tally() %>%
  ungroup() %>% 
  pivot_wider(id_cols = pgs_id,
              names_from = region,
              values_from = n, 
              names_prefix = "region_") %>% 
  mutate(total = !!sym(glue("region_{params$r1}")) + !!sym(glue("region_{params$r2}"))) %>% 
  filter(!is.na(total)) %>% 
  arrange(desc(total)) %>% 
  DT::datatable(options = list(pageLength = 10), rownames = FALSE, caption = "PGS with grand-progeny in both regions")

  
```

```{r, fig.width=7, fig.height=8}

pgsboth <- ped3 %>% group_by(pgs_id, region) %>% tally() %>% ungroup() %>% group_by(pgs_id) %>% filter(n_distinct(region) == 2) %>% ungroup() %>% distinct(pgs_id) %>% nrow()

ped3 %>%
  group_by(pgs_id, region) %>%
  tally() %>%
  ungroup() %>%
  group_by(pgs_id) %>%
  filter(n_distinct(region) == 2) %>%
  ungroup() %>% 
  ggplot(aes(x = n,
             fill = region)) +
  geom_histogram(bins = 20, 
                 show.legend = FALSE) +
  facet_wrap( ~ region, 
              nrow = 2, 
             labeller = labeller(
               region = c(
                 "1" = "1: Desert",
                 "2" = "2: Southeast",
                 "3" = "3: High Plains",
                 "4" = "4: Rainforest",
                 "5" = "5: Arid Prairie",
                 "6" = "6: Cold Desert",
                 "7" = "7: Forested Mountains",
                 "8" = "8: Fescue Belt",
                 "9" = "9: Upper Midwest & Northeast"
                 )
               )
              ) +
  scale_fill_manual(
    values = c(
      "1" = "tomato2",
      "2" = "darkslategray4",
      "3" = "springgreen3",
      "4" = "brown",
      "5" = "goldenrod1",
      "6" = "gray50",
      "7" = "deeppink3",
      "8" = "gray17",
      "9" = "slateblue2"
    )
  ) +
  labs(title = glue("Distribution of grand-progeny (PGSs with\n progeny in both regions, n = {pgsboth})"),
       x = "Number of calves",
       y = "Count of bulls") +
  theme_classic() +
    theme(
    plot.title = element_text(
      size = 20,
      face = "italic",
      margin = margin(t = 0, r = 0, b = 13, l = 0)
    ),
    axis.title = element_text(
      size = 16 
    ),
    axis.title.y = element_text(
      margin = margin(t = 0, r = 13, b = 0, l = 0)
    ),
    axis.title.x = element_text(
      margin = margin(t = 13, r = 0, b = 0, l = 0)
    ),
    axis.text = element_text(
      size = 14
    ),
    legend.text = element_text(
      size = 14
    ),
    strip.text.x = element_text(
      size = 14
    )
  )
```

