---
title: "Bivariate bootstrapping: weaning weight"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(dplyr)
library(tidyr)
library(glue)
library(stringr)
library(magrittr)
library(tibble)
library(purrr)
library(ggplot2)
library(rlang)
library(tidylog)


source(here::here("source_functions/melt_aireml.R"))
source(here::here("source_functions/calculate_heritability.R"))
source(here::here("source_functions/read_bootstrap_corrs.R"))
source(here::here("source_functions/region_key.R"))

```


# Notes & questions

# Setup

## `bootstrap_ww_start.R`

* Filter down `animal_regions` to form sampling pool
    + Retain only rows where `var` is equal to "weight"
    + Retain only rows where `trait` is equal to "ww"
    + Remove rows where `year` is before 1990
    + Remove rows where `n_animals` is fewer than 15
    + Remove single sire single dam contemporary groups
    
## `bootstrap_ww_sample.R`

* Using `sample_until.R`, selects zip codes to be retained
* Extracts animals in `start` matching selected zip codes
* Writes data file and pedigree file for the selected region and iteration 

## Resulting data

```{r, message=FALSE, warning=FALSE}

corr_iters <-
  tibble(iter = c(1:10)) %>% 
  expand(iter, region = c(1, 2, 5, 7, 8, 9)) %>% 
  mutate(data = 
           purrr::map2( .x = iter, .y = region,
             ~ read_bootstrap_corrs(iteration = .x, r1 = 3, r2 = .y)
             )
         ) %>% 
  unnest() %>% 
  left_join(
    region_key %>% 
      select(region = num,
             desc)
  ) %>% 
  mutate(run = as.character(glue("iter{iter}_3v{region}"))) %>% 
  select(
    iteration = iter,
    region = desc,
    effect_1,
    effect_2,
    corr,
    run
  )
```

```{r, message=FALSE}

bootstrap_h2s <-
  tibble(iter = c(1:10)) %>% 
  expand(iter, region = c(1, 2, 5, 7, 8, 9)) %>% 
  mutate(data = 
           purrr::map2(.x = iter, .y = region, ~ read_bootstrap_covs(iteration = .x, r1 = 3, r2 = .y))) %>% 
  unnest()

```

# Heritabilities

```{r, eval=FALSE}
bootstrap_h2s %>% 
  group_by(key) %>% 
  summarise(mean(`Direct h2`), sd(`Direct h2`), n = n())
  
``` 

# Genetic correlations

> Direct

```{r}

corr_iters %>% 
  filter_at(vars(contains("effect")), all_vars(str_detect(., "direct"))) %>% 
  mutate(region = glue("{region} (direct)")) %>% 
  group_by(region) %>% 
  summarise(
    min_corr = min(corr),
    mean_corr = mean(corr),
    max_cor = max(corr),
    sd_corr = sd(corr),
    n_iter = n(),
    range_corr = max(corr)-min(corr)) %>% 
  arrange(desc(range_corr))

```

> Maternal

```{r}
corr_iters %>% 
  filter_at(vars(contains("effect")), all_vars(str_detect(., "milk"))) %>% 
  mutate(region = glue("{region} (milk)")) %>% 
  group_by(region) %>% 
  summarise(
    min_corr = min(corr),
    mean_corr = mean(corr),
    max_cor = max(corr),
    sd_corr = sd(corr),
    n_iter = n(),
    range_corr = max(corr)-min(corr)) %>% 
  arrange(desc(range_corr))
```

```{r, eval=FALSE}
corr_iters %>% 
  filter_at(vars(contains("effect")), all_vars(str_detect(., "milk"))) %>% 
  mutate(region = glue("{region} (milk)")) %>% 
  select(-contains("effect")) %>% 
  View()
```


```{r, eval = FALSE}
corr_iters %>% 
  filter_at(vars(contains("effect")), all_vars(str_detect(., "direct"))) %>% 
  mutate(region = glue("{region} (direct)")) %>% 
  select(-contains("effect")) %>% 
  View()
```


# Correlated response

["Here the correlated response ($CR_Y$) is the response of the trait in the environment (Y) in which it is expected to perform, given selection in a different environment (X), while the direct response ($R_Y$) is for selection in the environment in which the organisms will ultimately be reared."](https://www.genetics.org/content/167/4/1529)

> Direct

```{r}
bootstrap_h2s %>% 
  mutate(which = if_else(key == "High Plains", "hp", "other")) %>% 
  select(which, h = `Direct h2`, run) %>% 
  mutate(h = sqrt(h)) %>% 
  pivot_wider(names_from = which,
              values_from = h,
              id_cols = run) %>% 
  left_join(
    corr_iters %>% 
      filter_at(vars(contains("effect")), all_vars(str_detect(., "direct"))) %>% 
      select(corr, run)
  ) %>% 
  mutate(cordir_res = ((hp/other)*corr)) %>% 
  arrange(cordir_res)

```

> Maternal

```{r}
bootstrap_h2s %>% 
  mutate(which = if_else(key == "High Plains", "hp", "other")) %>% 
  select(which, h = `Maternal h2`, run) %>% 
  mutate(h = sqrt(h)) %>% 
  pivot_wider(names_from = which,
              values_from = h,
              id_cols = run) %>% 
  left_join(
    corr_iters %>% 
      filter_at(vars(contains("effect")), all_vars(str_detect(., "milk"))) %>% 
      select(corr, run)
  ) %>% 
  mutate(cordir_res = ((hp/other)*corr)) %>% 
  arrange(cordir_res) 

```

# Commentary

