---
title: "Environmental variance"
author: "Harly Durbin"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(tidylog)
```

# Notes & questions

* I think I need to be looking at the residual not the CG solutions

# Setup

```{r}
cg_regions <- read_rds(here::here("data/derived_data/import_regions/cg_regions.rds"))
```

```{r}
animal_regions <- read_rds(here::here("data/derived_data/import_regions/animal_regions.rds"))
```

# Summarize changes in CG solution variance over time

## Total population 

```{r, fig.width=10, fig.height=5.4}
cg_regions %>%
  group_by(year) %>%
  summarise(mean_cg = mean(cg_sol),
            var_cg = var(cg_sol),
            sd_cg = sd(cg_sol),
            mad_cg = mad(cg_sol))
```

```{r, fig.width=10, fig.height=5.4}
cg_regions %>%
  group_by(year) %>%
  summarise(mean_cg = mean(cg_sol),
            sd_cg = sd(cg_sol),
            var_cg = var(cg_sol)) %>%
  ggplot(aes(x = year,
             y = var_cg)) +
  geom_line() +
  geom_smooth() +
  theme_classic() +
  theme(plot.title = element_text(size = 22,
                                  face = "italic",
                                  margin = margin(t = 0,
                                                  r = 0,
                                                  b = 13,
                                                  l = 0)),
        axis.title = element_text(size = 16),
        axis.title.y = element_text(margin = margin(t = 0,
                                                    r = 13,
                                                    b = 0,
                                                    l = 0)),
        axis.title.x = element_text(margin = margin(t = 13,
                                                    r = 0,
                                                    b = 0,
                                                    l = 0)),
        axis.text = element_text(size = 14),
        legend.position = "none") +
  labs(x = NULL,
       y = "Variance of CG solutions",
       color = NULL)
  
```

## Stratified by region

```{r, eval = FALSE}
cg_regions %>% 
  group_by(year, region) %>%
  summarise(mean_cg = mean(cg_sol),
            var_cg = var(cg_sol),
            sd_cg = sd(cg_sol),
            mad_cg = mad(cg_sol))
```

```{r, fig.width=16, fig.height=4}
cg_regions %>% 
  filter(year != 2019) %>% 
  group_by(year, region) %>%
  summarise(mean_cg = mean(cg_sol),
            var_cg = var(cg_sol),
            sd_cg = sd(cg_sol),
            mad_cg = mad(cg_sol)) %>% 
  ggplot(aes(x = year,
             y = var_cg)) +
  geom_line() +
  geom_smooth(aes(color = forcats::as_factor(region))) +
  scale_color_manual(values = c("1" = "tomato2",
                                "2" = "darkslategray4",
                                "3" = "springgreen3",
                                "5" = "goldenrod1",
                                "7" = "deeppink3",
                                "8" = "gray17",
                                "9" = "slateblue2"),
                     labels = c("1" = "1: Desert",
                                "2" = "2: Southeast",
                                "3" = "3: High Plains",
                                "5" = "5: Arid Prairie",
                                "7" = "7: Forested Mountains",
                                "8" = "8: Fescue Belt",
                                "9" = "9: Upper Midwest & Northeast")) +
  expand_limits(x = 2020) +
  scale_x_continuous(breaks = c(1990, 2000, 2010, 2020),
                     labels = c("1990", "2000", "2010", "2020")) +
  theme_classic() +
  theme(axis.title = element_text(size = 16),
        axis.title.y = element_text(margin = margin(t = 0,
                                                    r = 13,
                                                    b = 0,
                                                    l = 0)),
        axis.title.x = element_text(margin = margin(t = 13,
                                                    r = 0,
                                                    b = 0,
                                                    l = 0)),
        axis.text = element_text(size = 14),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        legend.position = "none") +
  labs(x = NULL,
       y = "Variance of CG solutions") +
  facet_wrap(~ region, nrow = 1)
  
```

# Model change in variance through years

* Jared: "Fit a linear model of the dependent variable vs. time. Get residuals. Either take the absolute value or square the residuals. Fit a second linear model in which you regress the squared residuals against time."

```{r}
year_cg_mod <-
  # Regress value on year
  lm(cg_sol ~ year, data = cg_regions) %>% 
  # Augment original data with residuals from model
  broom::augment() %>% 
  # Square residuals
  mutate(sq_resid = .resid^2) %>% 
  # Regress squared residuals on year
  lm(sq_resid ~ year, data = .)
```

```{r}
summary(year_cg_mod)
```

```{r}
anova(year_cg_mod)
```

# Coefficient of variation

* WL p. 573: "One technical comment before proceeding is that simple scale effects can also result in a change in the variance — if the coefficient of variation remains constant as its mean changes, then its variance must also change."

## Actual adjusted weaning weights

```{r}
animal_regions %>% 
  group_by(year) %>% 
  summarise(mean = mean(weight),
            CV = sqrt((var(weight))/(mean(weight))))
```

## Contemporary group solutions

```{r}
cg_regions %>% 
  summarise(mean = mean(value),
            CV = sqrt((var(value))/(mean(value))))
```
