---
title: "Drought monitor"
author: "Harly Durbin"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(tidyr)
library(dplyr)
library(magrittr)
library(purrr)
library(ggplot2)
library(tidylog)
```

# Notes and questions

* Data from 1/1/2000 to 5/1/2019 downloaded from USDA Drought Monitor 
    + [DSCI index](https://droughtmonitor.unl.edu/About/AbouttheData/DSCI.aspx)
* FIPS county code/ZIP4 cross-reference file came from https://www.kaggle.com/danofer/zipcodes-county-fips-crosswalk/version/1#
* [Grass-Cast](https://grasscast.unl.edu/)


# Setup 

```{r}
drought
```


```{r}
drought <-
  list.files(here::here("data/raw_data/drought_monitor"), full.names = TRUE) %>%
  purrr::map_dfr( ~ read_csv(.x)) %>%
  select(-MapDate,-StatisticFormatID) %>%
  distinct() %>%
  janitor::clean_names() %>% 
  # filter(fips %in% cg_drought$fips) %>% 
  mutate(week = lubridate::week(valid_start),
         year = lubridate::year(valid_start)) %>% 
  mutate(dsci = (1*d0)+(2*d1)+(3*d2)+(4*d3)+(5*d4))
```

```{r}

consec <-
  purrr::map(.x = c("1", "2", "3", "4"),
             function(x = .x) {
               df <-
                 drought %>%
                 select(fips,
                        state,
                        week,
                        year,
                        contains("valid"),
                        dsci,
                        # rename drought percentage column to dx
                        dx := !!rlang::sym(glue::glue("d{x}"))) %>%
                 # If at least 50% of the county is in drought at the x severity
                 # level, make inx column TRUE
                 mutate(inx := if_else(dx >= 0.5, TRUE, FALSE)) %>%
                 arrange(fips, year, week)
               
               # consecutive weeks in drought column based on fips and inx column
               df <-
                 transform(df, consecutivex = ave(inx, data.table::rleid(fips, inx), FUN = seq_along))
               
               df <-
                 df %>%
                 mutate(consecutivex = if_else(inx == FALSE, 0, as.double(consecutivex))) %>%
                 rename(
                   # Rename drought % column
                   !!rlang::sym(glue::glue("d{x}")) := dx,
                   # Rename in drought column
                   !!rlang::sym(glue::glue("in{x}")) := inx,
                   # Rename weeks in drought column
                   !!rlang::sym(glue::glue("c{x}")) := consecutivex
                 )
               
               return(df)
             }) %>%
  reduce(full_join)

```

```{r}

cg_drought <-
  read_rds(here::here("data/derived_data/cg_regions.rds")) %>% 
  filter(trait == "ww") %>%
  filter(!region %in% c(4, 6)) %>% 
  filter(n_animals >= 5) %>% 
  filter(year >= 2000) %>% 
  left_join(
    read_csv(here::here(
      "data/raw_data/ZIP-COUNTY-FIPS_2017-06.csv"
    )) %>%
      janitor::clean_names() %>%
      select(zip, fips = stcountyfp) %>% 
      mutate(zip = as.character(zip))
  ) %>% 
  mutate(week = lubridate::week(weigh_date)) %>% 
  left_join(consec, by = c("fips", "week", "year")) %>% 
  select(cg_new, weigh_date, value, contains("1"), contains("2"), contains("3"), contains("4"), dsci, n_animals, year, week, region, desc, state, fips, zip, lat, lng) %>% 
  # This is too liberal, need to figure out something better for zip codes in more than one county 
  group_by(cg_new) %>% 
  filter(n() > 1) %>% 
  arrange(desc(dsci)) %>% 
  slice(1) %>% 
  ungroup()
```

```{r}
cg_drought %>% 
  group_by(region) %>% 
  filter(!is.na(dsci)) %>% 
  summarise(cor(value, dsci))
```


# Correlate

```{r}
list(1, 2, 3, 5, 7, 8, 9) %>%
  set_names() %>%
  purrr::map(
    ~ cg_drought %>%
      filter(region == .x) %>%
      select(value, c1, c2, c3, c4) %>%
      corrr::correlate() %>%
      mutate(region = .x) %>% 
      slice(1) %>% 
      select(-rowname, -value) %>% 
      select(region, everything())
  ) %>% 
  reduce(bind_rows)


```

```{r}
cg_drought
```


```{r}
cg_drought %>% 
  select(value, d4, c)
```


# Commentary 

I want to use drought data to identify robust individuals.

Objective: do weaning weight contemporary group solutions fall in response to drought?
* Is there a threshold drought level when weaning weights (via contemporary group) fall?
* If change and threshold exist, are they different between regions?
* ...Between individuals (i.e., are certain individuals more robust to drought, be it directly or via their dam's effect etc.)?

Two problems with doing that.
* How to reduce the dimensionality of the drought dataset? I can calculate a drought severity index over a large time frame, but that sacrifices a lot of time-specific information
* What time points in calf's life to relate drought data? Maybe a calculate severity index at a couple timepoints then fit a multi-variate model?

Weekly drought data from January 1, 2000 to May, 1 2019 was downloaded from the USDA drought monitor for counties in the continental United States. FIPS county codes were then matched to breeder zip codes. In cases where a zip code resided in more than one county...

* CG solution for weaning weight contemporary groups after 2000
