---
title: "Environmental variance"
author: "Harly Durbin"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(tidylog)
```



# Notes & questions

* I think I need to be looking at the residual not the CG solutions

# Setup

```{r}
cg_regions <- read_rds(here::here("data/derived_data/cg_regions.rds"))
```

```{r}
animal_regions <- read_rds(here::here("data/derived_data/animal_regions.rds"))
```

# Plot 

## Total population 

```{r, fig.width=10, fig.height=5.4}
cg_regions %>%
  # At least 5 animals per CG
  filter(n_animals > 4 & trait == "ww" & var == "cg_sol") %>%
  # Very little data at most recent and earliest years
  filter(!year %in% c(1972, 2019)) %>%
  group_by(year) %>%
  summarise(mean_cg = mean(value),
            #var_cg = var(log(value))
            var_cg = var(value),
            sd_cg = sd(value)
            ) %>%
  ggplot(aes(x = year,
             y = mean_cg
             #y = value
             )) +
  geom_line() +
  geom_ribbon(aes(
    ymin = mean_cg - sd_cg,
    ymax = mean_cg + sd_cg
    ),
    alpha = 0.1
    ) +
  #geom_point() +
  #geom_smooth(aes(color = "red")) +
  theme_classic() +
  theme(
    plot.title = element_text(
      size = 22,
      face = "italic",
      margin = margin(
        t = 0,
        r = 0,
        b = 13,
        l = 0
      )
    ),
    axis.title = element_text(size = 16),
    axis.title.y = element_text(margin = margin(
      t = 0,
      r = 13,
      b = 0,
      l = 0
    )),
    axis.title.x = element_text(margin = margin(
      t = 13,
      r = 0,
      b = 0,
      l = 0
    )),
    axis.text = element_text(size = 14),
    legend.position = "none"
  ) +
  labs(
    x = NULL,
    y = "Variance of CG solutions",
    title = "Contemporary group variance has increased over time",
    color = NULL
  )
  
```

## Stratified by region

```{r, fig.width=10, fig.height=5.4}
cg_regions %>% 
  filter(n_animals > 4) %>% 
  filter(!year %in% c(1972, 2019)) %>% 
  group_by(year, region) %>% 
  summarise(mean_cg = mean(value),
            var_cg = var(value)) %>% 
  ggplot(aes(x = year,
             y = var_cg,
             color = forcats::as_factor(region))) +
  geom_point() +
  geom_smooth() +
        scale_color_manual(
        values = c(
          "1" = "tomato2",
          "2" = "darkslategray4",
          "3" = "springgreen3",
          "4" = "brown",
          "5" = "goldenrod1",
          "6" = "gray50",
          "7" = "deeppink3",
          "8" = "gray17",
          "9" = "slateblue2",
          "All regions" = "red"
        ),
        labels = c(
          "1" = "1: Desert",
          "2" = "2: Southeast",
          "3" = "3: High Plains",
          "4" = "4: Rainforest",
          "5" = "5: Arid Prairie",
          "6" = "6: Cold Desert",
          "7" = "7: Forested Mountains",
          "8" = "8: Fescue Belt",
          "9" = "9: Upper Midwest & Northeast",
          "All regions" = "National average"
        )
      ) +
  theme_classic() +
        theme(
        plot.title = element_text(
          size = 22,
          face = "italic",
          margin = margin(
            t = 0,
            r = 0,
            b = 13,
            l = 0
          )
        ),
        axis.title = element_text(size = 16),
        axis.title.y = element_text(margin = margin(
          t = 0,
          r = 13,
          b = 0,
          l = 0
        )),
        axis.title.x = element_text(margin = margin(
          t = 13,
          r = 0,
          b = 0,
          l = 0
        )),
        axis.text = element_text(size = 14)
      ) +
  labs(x = NULL,
       y = "Variance of CG solutions",
       title = "Contemporary group variance has increased over time") +
  facet_wrap(~ region)
  
```


# Model change in variance through years

* Jared: "Fit a linear model of the dependent variable vs. time. Get residuals. Either take the absolute value or square the residuals. Fit a second linear model in which you regress the squared residuals against time."

```{r}
year_cg_mod <-
  # Regress value on year
  lm(value ~ year, data = cg_regions) %>% 
  # Augment original data with residuals from model
  broom::augment() %>% 
  # Square residuals
  mutate(sq_resid = .resid^2) %>% 
  # Regress squared residuals on year
  lm(sq_resid ~ year, data = .)
```

```{r}
summary(year_cg_mod)
```

```{r}
anova(year_cg_mod)
```

## Coefficient of variation

* WL p. 573: "One technical comment before proceeding is that simple scale effects can also result in a change in the variance — if the coefficient of variation remains constant as its mean changes, then its variance must also change."

### Actual adjusted weaning weights

```{r}
animal_regions %>% 
  filter(n_animals > 4 & trait == "ww" & var == "weight") %>% 
  filter(!year %in% c(1972, 2019)) %>% 
  group_by(year) %>% 
  summarise(mean = mean(value),
            CV = sqrt((var(value))/(mean(value))))
```

### Contemporary group solutions

```{r}
cg_regions %>% 
  filter(n_animals > 4 & trait == "ww" & var == "cg_sol") %>% 
  filter(!year %in% c(1972, 2019)) %>% 
  group_by(year) %>% 
  summarise(mean = mean(value),
            CV = sqrt((var(value))/(mean(value))))
```

