---
title: "Gibbs"
author: "Harly Durbin"
date: "9/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

nohup psrecord "/usr/local/bin/airemlf90 renf90.par &> aireml.iter7.3v8.out" --log /home/agiintern/regions/log/psrecord/varcomp_ww/aireml/aireml.iter7.3v8.log --include-children --interval 5 &

# Unfinished 200,000 rounds

* 10,000 round burn-in
* Sample every 20 rounds

## Import "Monte Carlo error by time series"

* Assuming effect 0 corresponds to residual?
* Effective sample size: "At least > 10 is recommended. > 30 may be better."

```{r}
read_table2(here::here("data/derived_data/varcomp_ww/iter9/3v8/postout"), skip = 4, n_max = 15, col_names = c("position", "eff1", "eff2", "trt1", "trt2", "mce", "mean", "hdp", "effective_sample_size", "median", "mode"))
```


## Import "posterior standard deviation"

```{r}
read_table2(here::here("data/derived_data/varcomp_ww/iter9/3v8/postout"), skip = 27, n_max = 15, col_names = c("position", "eff1", "eff2", "trt1", "trt2", "psd", "mean", "psd_interval_lo", "psd_interval_hi", "convergence", "auto_lag1", "auto_lag10", "auto_lag50", "independent_batches")) 
```

## Import samples

"The meaning of each column is:

Column 1: the sequential number for saved samples.
Column 2: the actual round in which each sample was drawn.
Column 3: the number of parameters.
Column 4 or later: actual samples ordered by the position index.

```{r, message=FALSE, warning=FALSE}
gibbs200 <-
  c("iter4/3v9", "iter5/3v2", "iter8/3v2", "iter9/3v8", "iter9/3v9", "iter10/3v1") %>%
  purrr::set_names() %>%
  purrr::map_dfr(~ read_table2(here::here(glue::glue("data/derived_data/varcomp_ww/{.x}/postgibbs_samples")), col_names = FALSE), .id = "region") %>% 
  janitor::remove_empty(which = c("rows", "cols")) %>% 
  rename(sample = X1, round = X2)
  select(-X3, -X2) %>% 
  select(region:X18) %>% 
  tidyr::pivot_longer(cols = X4:X18) %>%
rename(sample = X1, variable = name)
```

## Plot samples

* Residual for position 15

```{r, fig.width=10, fig.height=10}
gibbs200 %>% 
  ggplot(aes(x = sample, y = value, color = variable)) +
  geom_line() +
  facet_wrap(~ region)

```

# Test with 30,000 rounds

* No burn in yet
* Sample every 20 rounds
---

## Import "Monte Carlo error by time series"

Assuming effect 0 corresponds to residual?

```{r}
read_table2(here::here("data/derived_data/varcomp_ww/iter2/3v1/postout"), skip = 4, n_max = 15, col_names = c("position", "eff1", "eff2", "trt1", "trt2", "mce", "mean", "hdp", "effective_sample_size", "median", "mode"))
```

## Import "posterior standard deviation"

```{r}
read_table2(here::here("data/derived_data/varcomp_ww/iter2/3v1/postout"), skip = 27, n_max = 15, col_names = c("position", "eff1", "eff2", "trt1", "trt2", "psd", "mean", "psd_interval_lo", "psd_interval_hi", "convergence", "auto_lag1", "auto_lag10", "auto_lag50", "independent_batches")) 
```

## Import samples

```{r, message=FALSE, warning=FALSE}
gibbs30 <-
  c("3v1", "3v2", "3v5", "3v7", "3v8", "3v9") %>%
  purrr::set_names() %>% 
  purrr::map_dfr(~ read_table2(here::here(glue::glue("data/derived_data/varcomp_ww/iter2/{.x}/postgibbs_samples")), col_names = FALSE), .id = "region") %>% 
  janitor::remove_empty(which = c("rows", "cols")) %>% 
  select(-X3, -X2) %>% 
  select(region:X18) %>% 
  tidyr::pivot_longer(cols = X4:X18) %>%
rename(sample = X1, variable = name)
```

## Plot samples

```{r, fig.width=10, fig.height=10}
gibbs30 %>% 
  ggplot(aes(x = sample, y = value, color = variable)) +
  geom_line() +
  facet_wrap(~ region)

```

```{r, fig.width=5, fig.height=5}
gibbs_samples %>% 
  filter(variable == "X9") %>% 
  filter(region == "3v1") %>% 
  ggplot(aes(x = value)) +
  geom_histogram()

```