---
title: "Trends over time"
author: "Harly Durbin"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(tidylog)
library(visdat)
library(grid)
library(gganimate)

source(here::here("source_functions/solutions_map.R"))
source(here::here("source_functions/solutions_line.R"))

```

# Notes & questions

# Setup

```{r}
cg_regions <- read_rds(here::here("data/derived_data/cg_regions.rds"))
```


```{r}
animal_regions <- read_rds(here::here("data/derived_data/animal_regions.rds"))
```

---



```{r, cache=TRUE, eval=FALSE}
cg_regions %>%
  filter(!region %in% c(4, 6)) %>%
  filter(trait == "ww" & var == "cg_sol") %>%
  filter(n_animals > 4) %>%
  group_by(year) %>%
  mutate(year_sd = sd(value),
         year_mean = mean(value),
         keep = case_when(
           value > year_mean + year_sd ~ "YES",
           value < year_mean - year_sd ~ "YES",
           TRUE ~ "NO"
         )) %>% 
  ungroup() %>% 
  filter(keep == "YES") %>% 
#  sample_frac(0.2) %>%
      ggplot(aes(
        x = lng,
        y = lat,
        color = "goldenrod",
        size = n_animals
      )) +
      geom_point(
        alpha = 0.3,
        #size = 0.5
        ) +
      scale_size(
        range = c(1,4)
      ) +
      usa +
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +
      coord_map("albers", lat0 = 39, lat1 = 45) +
      cowplot::theme_map() +
      labs(
        x = NULL,
        y = NULL,
        color = NULL,
        title = str_wrap( "Weaning weight CG solutions, low to high", width = 55)
      ) +
      #Set the "anchoring point" of the legend (bottom-left is 0,0; top-right is 1,1)
      #Put bottom-left corner of legend box in bottom-left corner of graph
      theme(
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        # Padding around the legend
        # top, right, bottom, left
        # legend.box.margin = margin(b = 0, r = 0.2, unit = "cm"),
        # Padding around the plot
        plot.margin = margin(
          t = 0.7,
          r = 0,
          b = 0.7,
          l = 1,
          unit = "cm"
        ),
        plot.title = element_text(
          # size = 56,
          # family =  "lato",
          size = 22,
          vjust = 6,
          face = "italic"
        )
      ) +
      # https://stackoverflow.com/questions/32656553/plot-legend-below-the-graphs-and-legend-title-above-the-legend-in-ggplot2
      guides(
        color = FALSE,
        size = FALSE
      ) +
  gganimate::transition_reveal(value) %>% 
  animate()


# anim_save(filename = here::here("figures/ww_solutions_appear.gif"), nframes = 48, fps = 2)
```


# Summary: year-to-year CG solutions by region {.tabset}

* "Your EPDs are not affected by weather/climate/environment"

## Weaning weight


### Actual adjusted weights

```{r, fig.width=10, fig.height=5.4}

animal_regions %>%
  filter(!region %in% c(4, 6)) %>%
  solutions_line(
    effect_var = "weight",
    trait_var = "ww",
    y_lab = "Mean adjusted weaning weight",
    plot_title = "Year-to-year weaning weights by region",
    national_avg = TRUE
    )



ggsave(here::here("figures/ww_pheno.solutions_line_year.png"), width = 10, height = 5.4)
```


### Contemporary group solutions

```{r, fig.width=10, fig.height=5.4}

cg_regions %>%
  filter(!region %in% c(4, 6)) %>%
  solutions_line(
    effect_var = "cg_sol",
    trait_var = "ww",
    y_lab = "Mean CG solution",
    plot_title = "Weaning weight contemporary group solutions reflect\nyear-to-year environmental trends",
    national_avg = TRUE
    ) +
  lims(y = c(440, 580))



ggsave(here::here("figures/ww_cg.solutions_line_year.png"), width = 10, height = 5.4)
```

### Breeding value solutions

```{r, fig.width=10, fig.height=5.4, cache=TRUE}

animal_regions %>%
  filter(!region %in% c(4, 6)) %>%
  solutions_line(
    effect_var = "bv_sol",
    trait_var = "ww",
    y_lab = "Mean breeding value",
    plot_title = "Weaning weight genetic trend: 1973-2019",
    national_avg = TRUE
    )



#ggsave(here::here("figures/ww_bv.solutions_line_year.png"), width = 10, height = 5.4)
```

### Maternal effect solutions

```{r, fig.width=10, fig.height=5.4, cache=TRUE}

animal_regions %>%
  filter(!region %in% c(4, 6)) %>%
  solutions_line(
    effect_var = "mat_sol",
    trait_var = "ww",
    y_lab = "Mean maternal effect solution",
    plot_title = "Milk genetic trend: 1973-2019",
    national_avg = TRUE
    )



#ggsave(here::here("figures/ww_mat.solutions_line_year.png"), width = 10, height = 5.4)
```



## Post-weaning gain

(Removed data from 1972: weird outlier, only one region with data)

### Contemporary group solutions

```{r, fig.width=10, fig.height=5.4}

cg_regions %>%
  filter(!region %in% c(4, 6)) %>%
  solutions_line(
    effect_var = "cg_sol",
    trait_var = "pwg",
    y_lab = "Mean CG solution",
    plot_title = "Post-weaning gain contemporary group solutions reflect\nyear-to-year environmental trends",
    national_avg = TRUE
    ) 

ggsave(here::here("figures/pwg_cg.solutions_line_year.png"), width = 10, height = 5.4)

```

### Actual adjusted weights

```{r, fig.width=10, fig.height=5.4, cache=TRUE}

animal_regions %>%
  filter(!region %in% c(4, 6)) %>%
  solutions_line(
    effect_var = "weight",
    trait_var = "pwg",
    y_lab = "Mean adjusted post-weaning gain",
    plot_title = "Year-to-year post-weaning gain by region", 
    national_avg = TRUE
    )



ggsave(here::here("figures/pwg_pheno.solutions_line_year.png"), width = 10, height = 5.4)
```


### Breeding value solutions

```{r, fig.width=10, fig.height=5.4}

animal_regions %>%
  filter(!region %in% c(4, 6)) %>%
  solutions_line(
    effect_var = "bv_sol",
    trait_var = "pwg",
    y_lab = "Mean breeing value",
    plot_title = "Post-weaning gain genetic trend: 1973-2019",
    national_avg = TRUE
    ) 

#ggsave(here::here("figures/pwg_bv.solutions_line_year.png"), width = 10, height = 5.4)

```

* What's up with 1972 region 5??

```{r}

cg_regions %>%
  filter(!region %in% c(4, 6)) %>%
      filter(trait == "pwg" & var == "cg_sol") %>%
      mutate(yr = lubridate::year(weigh_date),
             region = as.character(region)) %>%
      group_by(region, yr) %>%
  summarise(mean = mean(value)) %>% 
      bind_rows(
        cg_regions %>%
          filter(trait == "pwg" & var == "cg_sol") %>%
          mutate(yr = lubridate::year(weigh_date)) %>%
          group_by(yr) %>%
          summarise(mean = mean(value)) %>%
          mutate(region = "All regions")
      ) %>% 
  arrange(yr)
```

# Commentary
 
During routine genetic evaluation, contemporary group solutions are discarded, or at least un
