---
author: "Harly Durbin"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: hide
params:
  rundate: "190812"
  r1: "3"
  r2: "8"
title: "`r glue::glue('Bivariate weaning weight model variance components results (region {params$r1} vs. region {params$r2})')`"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(dplyr)
library(tidyr)
library(glue)
library(stringr)
library(magrittr)
library(purrr)
library(ggplot2)
library(rlang)
library(tidylog)

source(here::here("source_functions/melt_aireml.R"))
source(here::here("source_functions/gather_solutions.R"))


region_key <-
  tribble(~num, ~abbrv, ~desc,
          2, "SE", "Southeast",
          8, "FB", "Fescue Belt",
          3, "HP", "High Plains", 
          5, "AP", "Arid Prairie",
          7, "FM", "Forested Mountains", 
          1, "D", "Desert",
          9, "UMWNE", "Upper Midwest & Northeast")

r1_abbrv <- 
  region_key %>% 
  filter(num == params$r1) %>% 
  pull(abbrv)
  
r1_desc <-  
  region_key %>% 
  filter(num == params$r1) %>% 
  pull(desc)
 

r2_abbrv <- 
  region_key %>% 
  filter(num == params$r2) %>% 
  pull(abbrv)


r2_desc <-  
  region_key %>% 
  filter(num == params$r2) %>% 
  pull(desc)

options(scipen=999)
```


# Import/setup


```{r}
sol_univ <-
  gather_univ(
    rundate = params$rundate,
    r1 = params$r1,
    r2 = params$r2,
    growth_trait = "ww"
  )


```


```{r}
sol_biv <-
  gather_biv(
    rundate = params$rundate,
    r1 = params$r1,
    r2 = params$r2,
    growth_trait = "ww"
  )

```




```{r}
both <- read_table2(here::here(glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/ped.{r1_abbrv}_{r2_abbrv}.txt")), col_names = FALSE)

r1dat <- read_table2(here::here(glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/{r1_abbrv}/data.{r1_abbrv}.txt")), col_names = FALSE)

r2dat <-  read_table2(here::here(glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/{r2_abbrv}/data.{r2_abbrv}.txt")), col_names = FALSE)

```


# Notes

## Univariate and bivariate full

* `r scales::comma(nrow(both))` total animals, with and without records
* `r scales::comma(nrow(r1dat))` from `r r1_desc` with records
* `r scales::comma(nrow(r2dat))` from `r r2_desc` with records

# Distribution of breeding values

## Univariate

```{r, fig.width=8, fig.height=6}

sol_univ %>%
  filter(effect == "bv_sol") %>%
  ggplot(aes(x = solution,
             fill = forcats::as_factor(region))) +
  geom_density(show.legend = FALSE) +
  facet_wrap(~ region,
             nrow = 2,
             labeller = labeller(
               region = c(
                 "1" = "1: Desert",
                 "2" = "2: Southeast",
                 "3" = "3: High Plains",
                 "4" = "4: Rainforest",
                 "5" = "5: Arid Prairie",
                 "6" = "6: Cold Desert",
                 "7" = "7: Forested Mountains",
                 "8" = "8: Fescue Belt",
                 "9" = "9: Upper Midwest & Northeast"
                 )
               )
             ) +
  scale_fill_manual(
    values = c(
      "1" = "tomato2",
      "2" = "darkslategray4",
      "3" = "springgreen3",
      "4" = "brown",
      "5" = "goldenrod1",
      "6" = "gray50",
      "7" = "deeppink3",
      "8" = "gray17",
      "9" = "slateblue2"
    )
  ) +
  labs(title = "Distribution of breeding values, single-trait model",
       x = "Breeding value",
       y = "Kernel density") +
  theme_classic() +
    theme(
    plot.title = element_text(
      size = 20,
      face = "italic",
      margin = margin(t = 0, r = 0, b = 13, l = 0)
    ),
    axis.title = element_text(
      size = 16 
    ),
    axis.title.y = element_text(
      margin = margin(t = 0, r = 13, b = 0, l = 0)
    ),
    axis.title.x = element_text(
      margin = margin(t = 13, r = 0, b = 0, l = 0)
    ),
    axis.text = element_text(
      size = 14
    ),
    legend.text = element_text(
      size = 14
    ),
    strip.text.x = element_text(
      size = 14
    )
  )

```

## Bivariate

```{r, fig.width=8, fig.height=6}

sol_biv %>% 
  filter(effect == "bv_sol") %>% 
  ggplot(aes(
    x = solution,
    fill = forcats::as_factor(region)
  )) +
  geom_density(show.legend = FALSE) +
  facet_wrap( ~ region, 
              nrow = 2, 
              labeller = labeller(
               region = c(
                 "1" = "1: Desert",
                 "2" = "2: Southeast",
                 "3" = "3: High Plains",
                 "4" = "4: Rainforest",
                 "5" = "5: Arid Prairie",
                 "6" = "6: Cold Desert",
                 "7" = "7: Forested Mountains",
                 "8" = "8: Fescue Belt",
                 "9" = "9: Upper Midwest & Northeast"
                 )
               )
              ) +
  scale_fill_manual(
    values = c(
      "1" = "tomato2",
      "2" = "darkslategray4",
      "3" = "springgreen3",
      "4" = "brown",
      "5" = "goldenrod1",
      "6" = "gray50",
      "7" = "deeppink3",
      "8" = "gray17",
      "9" = "slateblue2"
    )
  ) +
  labs(title = "Distribution of breeding values, multi-trait model",
       x = "Breeding value",
       y = "Kernel density") +
  theme_classic() +
    theme(
    plot.title = element_text(
      size = 20,
      face = "italic",
      margin = margin(t = 0, r = 0, b = 13, l = 0)
    ),
    axis.title = element_text(
      size = 16 
    ),
    axis.title.y = element_text(
      margin = margin(t = 0, r = 13, b = 0, l = 0)
    ),
    axis.title.x = element_text(
      margin = margin(t = 13, r = 0, b = 0, l = 0)
    ),
    axis.text = element_text(
      size = 14
    ),
    legend.text = element_text(
      size = 14
    ),
    strip.text.x = element_text(
      size = 14
    )
  )

```


# Heritability

## Univariate

### `r r1_desc`

```{r, eval = TRUE}

melt_aireml(
  path = glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/{r1_abbrv}/airemlf90.{r1_abbrv}.log"),
  effect2 = c(glue("{r1_abbrv}_dir"), glue("{r1_abbrv}_mat")),
  effect4 = c(glue("{r1_abbrv}_mpe")),
  resids = c(glue("{r1_abbrv}_resid"))
  ) %>% 
  mutate_at(vars(contains("val")), ~ str_remove(., glue("{r1_abbrv}_"))) %>%
  mutate(
    val1 = 
      case_when(
        val1 == "dir" & val2 == "mat" ~ "dir_mat",
        TRUE ~ val1
        ),
    val2 = if_else(val1 == "dir_mat", val1, val2)
    ) %>% 
  filter(val1 == val2) %>% 
  select(-val2) %>% 
  spread(val1, var_cov) %>% 
  summarise(
    !!glue("Direct h2 ({ensym(r1_desc)})") := dir/sum(dir, mat, dir_mat, mpe, resid),
    !!glue("Milk h2 ({ensym(r1_desc)})") := mat/sum(dir, mat, dir_mat, mpe, resid),
    !!glue("MPE c2 ({ensym(r1_desc)})") := mpe/sum(dir, mat, dir_mat, mpe, resid),
    `Total variance` = sum(dir, mat, dir_mat, mpe, resid)
    ) 

```

### `r r2_desc`

```{r, eval = TRUE}

melt_aireml(
  path = glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/{r2_abbrv}/airemlf90.{r2_abbrv}.log"),
  effect2 = c(glue("{r2_abbrv}_dir"), glue("{r2_abbrv}_mat")),
  effect4 = c(glue("{r2_abbrv}_mpe")),
  resids = c(glue("{r2_abbrv}_resid"))
  ) %>% 
  mutate_at(vars(contains("val")), ~ str_remove(., glue("{r2_abbrv}_"))) %>%
  mutate(
    val1 = 
      case_when(
        val1 == "dir" & val2 == "mat" ~ "dir_mat",
        TRUE ~ val1
        ),
    val2 = if_else(val1 == "dir_mat", val1, val2)
    ) %>% 
  filter(val1 == val2) %>% 
  select(-val2) %>% 
  spread(val1, var_cov) %>% 
  summarise(
    !!glue("Direct h2 ({ensym(r2_desc)})") := dir/sum(dir, mat, dir_mat, mpe, resid),
    !!glue("Milk h2 ({ensym(r2_desc)})") := mat/sum(dir, mat, dir_mat, mpe, resid),
    !!glue("MPE c2 ({ensym(r2_desc)})") := mpe/sum(dir, mat, dir_mat, mpe, resid),
    `Total variance` = sum(dir, mat, dir_mat, mpe, resid)
    ) 

```




## Bivariate

```{r}
melt_aireml(
  path = glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/airemlf90.{r1_abbrv}_{r2_abbrv}.log"),
  effect2 = c(glue("{r1_abbrv}_dir"), glue("{r2_abbrv}_dir"), glue("{r1_abbrv}_mat"), glue("{r2_abbrv}_mat")),
  effect4 = c(glue("{r1_abbrv}_mpe"), glue("{r2_abbrv}_mpe")),
  resids = c(glue("{r1_abbrv}_resid"), glue("{r2_abbrv}_resid"))
  ) %>% 
  mutate(key = 
           case_when(
             str_detect(val1, r1_abbrv) ~ r1_desc, 
             str_detect(val1, r2_abbrv) ~ r2_desc,
             ),
         val1 = 
           case_when(
             val1 == glue("{r1_abbrv}_dir") & val2 == glue("{r1_abbrv}_mat") ~ glue("{r1_abbrv}_dir_mat"),
             val1 == glue("{r2_abbrv}_dir") & val2 == glue("{r2_abbrv}_mat") ~glue("{r2_abbrv}_dir_mat"),
             TRUE ~ val1
           ),
         val2 = if_else(str_detect(val1, "dir_mat"), val1, val2)
         ) %>% 
  filter(val1 == val2) %>% 
  select(-val2) %>% 
  mutate(val1 = str_remove(val1, glue("{r1_abbrv}_|{r2_abbrv}_"))) %>% 
  tidyr::pivot_wider(
    id_cols = key,
    names_from = val1,
    values_from = var_cov
    ) %>% 
  group_by(key) %>% 
  summarise(
    !!glue("Direct h2") := dir/sum(dir, mat, dir_mat, mpe, resid),
    !!glue("Milk h2") := mat/sum(dir, mat, dir_mat, mpe, resid),
    !!glue("MPE c2") := mpe/sum(dir, mat, dir_mat, mpe, resid),
    `Total variance` = sum(dir, mat, dir_mat, mpe, resid)
    )
```


# Genetic correlations

## Full bivariate run

### WW direct & milk

```{r}
read_table2(here::here(glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/airemlf90.{r1_abbrv}_{r2_abbrv}.log")),
            skip = 9,
            n_max = 4,
            col_names = c(glue("{r1_desc} (direct)"), glue("{r2_desc} (direct)"), glue("{r1_desc} (milk)"), glue("{r2_desc} (milk)"))) %>%
  mutate(rowname = colnames(.)) %>% 
  select(rowname, everything())
```

### Maternal permanent environment

```{r}
read_table2(here::here(glue("data/f90/{params$rundate}_{params$r1}v{params$r2}_ww/airemlf90.{r1_abbrv}_{r2_abbrv}.log")),
            skip = 24,
            n_max = 2,
 col_names = c(glue("{r1_desc} (MPE)"), glue("{r2_desc} (MPE)"))) %>%
  mutate(rowname = colnames(.)) %>% 
  select(rowname, everything())
```

